Paper 1

### 2022-06-15 scripts/Paper 1.Rmd
SWC plot amend

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(plotly)
library(plyr)
library(dplyr)
library(knitr)
library(tidyverse)
#library(devtools)
library(openair)
library(tidyr)
library(naniar)
library(forecast)
library(kableExtra)
library(htmltools)
library(ggpubr)
library(rstatix)
library(broom)
library(gtable)
library(gridExtra)
library(cowplot)

@@

input.daily$month <- as.numeric(input.daily$month) 
input.daily$month <- month.abb[input.daily$month]

@@

input.rearranged <- rbind(Grow19.20,Grow20.21, Grow21.22)
```

@@

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
(SP.rain <- input.daily %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
   scale_x_discrete(limits = month.abb) +
  theme_classic()+
  theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-4]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[1] <- "jday"
colnames(rainfall_L)[3] <- "month"
rainfall_L$Site <- "L"
rainfall_L <- rainfall_L %>% filter(month >= 6)
rainfall_L$month <- as.numeric(rainfall_L$month) 
rainfall_L$month<- month.abb[rainfall_L$month]
rainfall_L$GSYear <- "19/20"

@@

# define seasons
winter <- (rainfall_L$month == "Jun" | rainfall_L$month == "Jul" | rainfall_L$month == "Aug")
spring <- (rainfall_L$month == "Sep" | rainfall_L$month == "Oct" | rainfall_L$month == "Nov")
earlysummer <- (rainfall_L$month == "Dec")
latesummer <- (rainfall_L$month == "Jan" | rainfall_L$month == "Feb")
autumn <- (rainfall_L$month == "Mar" | rainfall_L$month == "Apr" | rainfall_L$month == "May")

rainfall_L$season[winter] <- "Winter"
rainfall_L$season[spring] <- "Spring"
rainfall_L$season[earlysummer] <- "EarlySummer"
rainfall_L$season[latesummer] <- "LateSummer"
rainfall_L$season[autumn] <- "Autumn"

rainfall_L$season <- factor(rainfall_L$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))

rainfall_L %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("month")+
  theme_classic()

rainfall.SP <- input.daily %>% select(Precip, jday, month, Year)
rainfall.SP$Site <- "SP"
rainfall <- rbind(rainfall_L, rainfall.SP)

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))

@@

Mean daily air temperature (top) and monthly (bottom)
Mean daily air temperature

```{r Air temperature, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
TA <- ggplot(input.daily, aes(x = jday, y = Ta.mean, group = Year)) +
  geom_line(aes(color = Year), size =1)+
    xlab("Julian day of year")+
    ylab("Air temperature (°C)")+
     #geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))

(TS <- ggplot(input.daily, aes(x = jday, y = Ts.mean, group = Year)) +
  geom_line(aes(color = Year), size =1)+
    xlab("Julian day of year")+
    ylab("Soil temperature (°C)")+
     #geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

ggplotly(ggplot(input.daily, aes(x = jday, y = Ta.max, group = Year)) +
  geom_line(aes(color = Year))+
  theme_classic())

ggplotly(ggplot(input.daily, aes(x = jday, y = Ta.min, group = Year)) +
  geom_line(aes(color = Year))+
  theme_classic())
```

@@

```{r, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
(SWC <- ggplot(input.daily, aes(jday, SWC, group = Year)) +
  geom_line(aes(color = Year), size = 1.5)+
  xlab('Julian day of year') +
  ylab('SWC %') +
  theme_classic() + 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))
```

@@

Fig1 <- plot_grid(SP.rain + theme(legend.position="none"), TA+ theme(legend.position="none"),
             SWC+ theme(legend.position="none"), TS+ theme(legend.position="none"), ncol = 2, rel_heights = c(0.5, 0.5, 0.5, 0.5))

@@

```{r add seasons, include=FALSE}
rainfall$season <- "" # add season variable

# define seasons
winter <- (rainfall$month == "Jun" | rainfall$month == "Jul" | rainfall$month == "Aug")
spring <- (rainfall$month == "Sep" | rainfall$month == "Oct" | rainfall$month == "Nov")
earlysummer <- (rainfall$month == "Dec")
latesummer <- (rainfall$month == "Jan" | rainfall$month == "Feb")
autumn <- (rainfall$month == "Mar" | rainfall$month == "Apr" | rainfall$month == "May")

rainfall$season[winter] <- "Winter"
rainfall$season[spring] <- "Spring"
rainfall$season[earlysummer] <- "EarlySummer"
rainfall$season[latesummer] <- "LateSummer"
rainfall$season[autumn] <- "Autumn"

rainfall$season <- factor(rainfall$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))
```

@@

```{r cumulative fluxes, fig.width = 10, fig.height = 6}
cum.co2fluxes <- input.rearranged %>% #input.daily %>%
  select(NEE_LT.gC, GPP_LT.gC, ER_LT.gC, dayssinceJul1, GSYear) %>% 
  group_by(GSYear) %>% 
  mutate(NEEcum = cumsum(NEE_LT.gC),
         GPPcum = cumsum(GPP_LT.gC),
         ERcum = cumsum(ER_LT.gC))

@@

# define seasons
winter <- (input.rearranged$month == "Jun" | input.rearranged$month == "Jul" | input.rearranged$month == "Aug")
spring <- (input.rearranged$month == "Sep" | input.rearranged$month == "Oct" | input.rearranged$month == "Nov")
earlysummer <- (input.rearranged$month == "Dec")
latesummer <- (input.rearranged$month == "Jan" | input.rearranged$month == "Feb")
autumn <- (input.rearranged$month == "Mar" | input.rearranged$month == "Apr" | input.rearranged$month == "May")

input.rearranged$season[winter] <- "Winter"
input.rearranged$season[spring] <- "Spring"
input.rearranged$season[earlysummer] <- "EarlySummer"
input.rearranged$season[latesummer] <- "LateSummer"
input.rearranged$season[autumn] <- "Autumn"

input.rearranged$season <- factor(input.rearranged$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))
```

```{r}
rainfall.SP <- input.rearranged %>% select(Precip, month, GSYear, season)

rainfall <- rainfall_L %>% select(Precip, month, GSYear, season) %>% rbind(., rainfall.SP)

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("GSYear", "season"), summarise,
  seasonal.total = sum(Precip)) %>%
  ggplot(., aes(x = GSYear, y = seasonal.total, group = GSYear, fill = season)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col() + 
  ylab("Seasonal rainfall (mm)")+
  xlab("Month")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))
```

>>>>

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(plotly)
library(plyr)
library(dplyr)
library(knitr)
library(tidyverse)
#library(devtools)
library(openair)
library(tidyr)
library(naniar)
library(forecast)
library(kableExtra)
library(htmltools)
library(ggpubr)
library(rstatix)
library(broom)
library(gtable)
library(gridExtra)
library(cowplot)
library(viridis)
library(zoo)

@@

input.daily$month <- as.numeric(input.daily$month) 
input.daily$month.name <- month.abb[input.daily$month]
input.daily$date <- as.Date(with(input.daily,paste(Year,month,jday,sep="-")),"%Y-%m-%j")

@@

input.rearranged <- rbind(Grow19.20,Grow20.21, Grow21.22)
input.rearranged$monthyear <- as.yearmon(paste(input.rearranged$month, input.rearranged$Year), "%m %Y")

start.time <- as.Date("07-01", "%m-%d")
end.time <- as.Date("06-30", "%m-%d")
start.end <- c(start.time, end.time)
```

@@

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-4]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[1] <- "jday"
colnames(rainfall_L)[3] <- "month"
rainfall_L$Site <- "L"
rainfall_L <- rainfall_L %>% filter(month >= 7)
rainfall_L$month <- as.numeric(rainfall_L$month) 
rainfall_L$month.name<- month.abb[rainfall_L$month]
rainfall_L$GSYear <- "19/20"

@@

# define seasons
winter <- (rainfall_L$month.name == "Jun" | rainfall_L$month.name == "Jul" | rainfall_L$month.name == "Aug")
spring <- (rainfall_L$month.name == "Sep" | rainfall_L$month.name == "Oct" | rainfall_L$month.name == "Nov")
summer <- (rainfall_L$month.name == "Dec" | rainfall_L$month.name == "Jan" | rainfall_L$month.name == "Feb")
autumn <- (rainfall_L$month.name == "Mar" | rainfall_L$month.name == "Apr" | rainfall_L$month.name == "May")

rainfall_L$season[winter] <- "Winter"
rainfall_L$season[spring] <- "Spring"
rainfall_L$season[summer] <- "Summer"
rainfall_L$season[autumn] <- "Autumn"

rainfall_L$season <- factor(rainfall_L$season , levels=c('Summer', 'Autumn', 'Winter', 'Spring'))
```

```{r, fig.width=11, fig.height=4}
rainfall.SP <- input.rearranged %>% select(Precip, month, GSYear, season, Year, month.name)

rainfall <- rainfall_L %>% select(Precip, month, GSYear, season, Year, month.name) %>% rbind(., rainfall.SP)

rainfall$monthyear <- as.yearmon(paste(rainfall$month, rainfall$Year), "%m %Y")

(rain <- rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("GSYear", "season"), summarise,
  seasonal.total = sum(Precip)) %>%
  ggplot(., aes(x = GSYear, y = seasonal.total, group = GSYear, fill = season)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col() + 
  ylab("Seasonal rainfall (mm)")+
  xlab("Growing season year")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

(rainfall.p <-rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("monthyear", "Year"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = monthyear, y = monthly.total)) +
  geom_col(position = "dodge") +
  scale_x_yearmon(n = 7, expand = c(0,0))+
  ylab("Monthly rainfall (mm)")+
  xlab("Date")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14),
          axis.text.x = element_text( vjust = 0.5)))
rainfall.p
tiff("rainfall.tiff", units="in", width=10, height=6,  res=300)
print(rainfall.p)
dev.off()

@@

```{r Air temperature, echo=FALSE, message = FALSE, fig.width = 11, fig.height = 4}
(TA <- ggplot(input.rearranged, aes(x = date, y = Ta.mean)) +
   geom_point(size = 0.001)+
   geom_line(size = 0.72)+
        geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
    xlab("Date")+
    ylab("Air temperature (°C)")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

(TS <- ggplot(input.rearranged, aes(x = date, y = Ts.mean)) +
   geom_point(size = 0.001)+
   geom_line(size = 0.72)+
        geom_ribbon(aes(ymin=Ts.min, ymax=Ts.max), alpha=.3, linetype=0) +
    xlab("Date")+
    ylab("Soil temperature (°C)")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

TA
tiff("TA.tiff", units="in", width=10, height=6,  res=300)
print(TA)
dev.off()

TS
tiff("TS.tiff", units="in", width=10, height=6,  res=300)
print(TS)
dev.off()
```

@@

```{r, echo=FALSE, message = FALSE, fig.width = 7, fig.height = 4}
group.colors <- c("19/20" = "#333BFF", "20/21" = "#CC6600", "21/22" ="#9633FF")

(SWC.20 <- input.rearranged %>% filter (GSYear == "19/20"| GSYear == "20/21" | GSYear == "21/22" ) %>% ggplot(., aes(x = as.Date(dayssinceJul1, origin = as.Date("07-01", "%m-%d")), SWC, group = GSYear)) +
  geom_line(aes(color = GSYear), size = 1.5)+
  # scale_x_continuous(limits = c(0, 365))+
  xlab('Month') +
  ylab('SWC %') +
   scale_x_date(breaks = "month", date_labels = "%b", limits = as.Date(c("07-01", "06-31"), "%m-%d"))+
    scale_color_manual(values = group.colors)+
  theme_classic() + 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

SWC.20
tiff("SWC22.tiff", units="in", width=10, height=6,  res=300)
print(SWC.20)
dev.off()
```

@@

Fig1 <- plot_grid(rainfall.p + theme(legend.position="none"), TA+ theme(legend.position="none"),
             SWC.20+ theme(legend.position="none"), TS+ theme(legend.position="none"), ncol = 2, rel_heights = c(0.5, 0.5, 0.5, 0.5))

@@

```{r add seasons, include=FALSE}
#rainfall$season <- "" # add season variable

# define seasons
#winter <- (rainfall$month == "Jun" | rainfall$month == "Jul" | rainfall$month == "Aug")
#spring <- (rainfall$month == "Sep" | rainfall$month == "Oct" | rainfall$month == "Nov")
#summer <- (rainfall$month == "Dec" | rainfall$month == "Jan" | rainfall$month == "Feb")
#autumn <- (rainfall$month == "Mar" | rainfall$month == "Apr" | rainfall$month == "May")

#rainfall$season[winter] <- "Winter"
#rainfall$season[spring] <- "Spring"
#rainfall$season[summer] <- "Summer"
#rainfall$season[earlysummer] <- "EarlySummer"
#rainfall$season[latesummer] <- "LateSummer"
#rainfall$season[autumn] <- "Autumn"

#rainfall$season <- factor(rainfall$season , levels=c('Summer', 'Autumn', 'Winter', 'Spring'))
```

@@

```{r cumulative fluxes, fig.width = 10, fig.height = 6}
cum.co2fluxes <- input.rearranged %>% #input.daily %>%
  select(NEE_LT.gC, GPP_LT.gC, ER_LT.gC, dayssinceJul1, GSYear, season) %>% 
  group_by(GSYear) %>% 
  mutate(NEEcum = cumsum(NEE_LT.gC),
         GPPcum = cumsum(GPP_LT.gC),
         ERcum = cumsum(ER_LT.gC))

@@

# define seasons
winter <- (input.rearranged$month.name == "Jun" | input.rearranged$month.name == "Jul" | input.rearranged$month.name == "Aug")
spring <- (input.rearranged$month.name == "Sep" | input.rearranged$month.name == "Oct" | input.rearranged$month.name == "Nov")
summer <- (input.rearranged$month.name == "Dec" | input.rearranged$month.name == "Jan" | input.rearranged$month.name == "Feb")
autumn <- (input.rearranged$month.name == "Mar" | input.rearranged$month.name == "Apr" | input.rearranged$month.name == "May")

input.rearranged$season[winter] <- "Winter"
input.rearranged$season[spring] <- "Spring"
input.rearranged$season[summer] <- "Summer"
#input.rearranged$season[latesummer] <- "LateSummer"
input.rearranged$season[autumn] <- "Autumn"

input.rearranged$season <- factor(input.rearranged$season , levels=c('Summer', 'Autumn', 'Winter', 'Spring'))
```

```{r}
(GPPcum.p <-cum.co2fluxes %>% 
   ddply(c("GSYear", "season"), summarise,
  seasonal.total = cumsum(GPP_LT.gC)) %>%
   ggplot(., aes(x = GSYear, y = seasonal.total, group = GSYear, fill = season)) +
  geom_col() + 
  ylab("Total GPP")+
  xlab("Growing season year")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

(ERcum.p <-cum.co2fluxes %>% 
   ddply(c("GSYear", "season"), summarise,
  seasonal.total = cumsum(ER_LT.gC)) %>%
   ggplot(., aes(x = GSYear, y = seasonal.total, group = GSYear, fill = season)) +
  geom_col() + 
  ylab("Total ER")+
  xlab("Growing season year")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))
```

```{r, fig.height=10, fig.width = 6}

(Fig3 <- plot_grid(rain,
             GPPcum.p, ERcum.p, nrow =3))
